---
title: '<center> <h1> A taste of data science using R and the ```tidyverse``` -- Solutions </h1> </center>'
author: '[Antonio Schettino](https://asch3tti.netlify.com/ "Antonio Schettino")'
date: 'January 15th, 2020'
output:
  html_document:
    df_print: paged
    theme: spacelab
    highlight: tango
    code_folding: show
    toc: true
    toc_float: true
    includes:
      after_body: footer.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

library(here)
library(tidyverse)
library(yarrr)

```

# Solutions

## Exercise 1

Install and load the ```tidyverse```.

```{r install_load_tidyverse, eval = FALSE}

install.packages("tidyverse")
library(tidyverse)

```

## Exercise 2

Do the following operations _**in one go**_:

* load the original data (```MixedAttitude.dat```)
* convert ```sex``` to a **categorical** variable (named ```gender```) with 2 levels:
    - 1 --> female
    - 2 --> male
* eliminate from the data set the following variables: ```sex```, ```beerpos```, ```beerneg```, ```beerneut```, ```winepos```, and ```waterpos```
* using the function ```rename```, rename the variables you kept:
    - ssj --> participant
    - wineneg --> wine_negative
    - wineneut --> wine_neutral
    - waterneg --> water_negative
    - waterneu --> water_neutral
* filter out outlier participants as well as all participants who rated _water_ after _neutral_ imagery as _**lower than -10**_
* convert the data set to **long** format
* separate conditions into 2 variables (```drink``` and ```imagery```)
* convert ```participant```, ```gender```, ```drink```, and ```imagery``` to factors
* save the data set as ```data_attitude_wine.csv``` in the subfolder ```tidy-data```
* separately for ```gender```, ```drink```, and ```imagery```, calculate the following summary statistics:
    - number of observations
    - median
    - median absolute deviation
    - minimum value
    - maximum value
* display the results in console

```{r exercise_2}

data_attitude_wine <-
  read_csv(
    here(
      "raw-data",
      "MixedAttitude.dat"
    )
  ) %>%
  mutate(
    gender = recode(
      factor(sex),
      "1" = "female", "2" = "male"
    )
  ) %>%
  select(-c(
    sex,
    beerpos,
    beerneg,
    beerneut,
    winepos,
    waterpos)) %>%
  rename(participant = ssj,
         wine_negative = wineneg,
         wine_neutral = wineneut,
         water_negative = waterneg,
         water_neutral = waterneu) %>%
  filter(
    water_neutral >= -100 & water_neutral <= +100 &
      water_neutral >= -10
  ) %>%
  pivot_longer(
    cols = c(wine_negative, wine_neutral, water_negative, water_neutral),
    names_to = "condition",
    values_to = "ratings",
  ) %>%
   separate(
    condition, c("drink", "imagery"),
    remove = TRUE
  ) %>%
    mutate(
    participant = as_factor(participant),
    gender = as_factor(gender),
    drink = as_factor(drink),
    imagery = as_factor(imagery)
  )

write_csv(
  data_attitude_wine,
  here("tidy-data", "data_attitude_wine.csv")
)

summary_data_attitude_wine <-
data_attitude_wine %>%
    group_by(gender, drink, imagery) %>%
    summarize(n = n(),
              median = median(ratings),
              mad = mad(ratings),
              min = min(ratings),
              max = max(ratings)) %>%
    print()

```

## Exercise 3

Load ```data_attitude_wine.csv```.

```{r exercise_3_0}

data_attitude_wine <-
  read_csv(
    here(
      "tidy-data",
      "data_attitude_wine.csv"
    )
  )

```

1. create a plot similar to what is shown above (with ```wine``` instead of ```beer```) and save it in the subfolder ```figures```

```{r exercise_3_1}

RDIplot_ex_3_1 <-
  data_attitude_wine %>%
  ggplot(aes(
    x = imagery,
    y = ratings,
    color = drink,
    fill = drink
  )) +
  geom_pirate(
    bars = FALSE,
    cis = TRUE,
    lines = TRUE, lines_params = list(color = "black"),
    points = TRUE, points_params = list(shape = 21, color = "black", size = 5, alpha = .4),
    violins = TRUE, violins_params = list(size = 1),
    show.legend = TRUE
  ) +
  scale_y_continuous(
    limits = c(-100, 100),
    breaks = seq(-100, 100, 10)
  ) +
  coord_cartesian(ylim = c(-40, 40)) +
  geom_hline(
    yintercept = seq(-100, 100, 10),
    linetype = "dotted",
    colour = "#999999",
    size = .8,
    alpha = .5
  ) +
  scale_fill_viridis(
    option = "viridis",
    discrete = TRUE
  ) +
  scale_color_viridis(
    option = "viridis",
    discrete = TRUE
  ) +
  ggtitle("Plot Exercise 3.1") +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid = element_blank(),
    legend.box.background = element_rect(color = "transparent"),
    legend.position = c(.9, .9),
    plot.title = element_text(size = 26, hjust = .5)
  )

RDIplot_ex_3_1

ggsave(filename = "plot_ex_3_1.png",
       plot = RDIplot_ex_3_1,
       path = here("doc/images"),
       width = 7, height = 7, units = "in",
       dpi = 600)

```

2. the same as above, separately for female and male participants (**hint**: use ```facet_wrap```)

```{r exercise_3_2}

RDIplot_ex_3_2 <-
  RDIplot_ex_3_1 +
  ggtitle("Plot Exercise 3.2") +
  facet_wrap(. ~ gender)

RDIplot_ex_3_2

ggsave(filename = "plot_ex_3_2.png",
       plot = RDIplot_ex_3_2,
       path = here("doc/images"),
       width = 9, height = 7, units = "in",
       dpi = 600)

```

```{r exercise_3_bonus}

# following https://bookdown.org/ndphillips/YaRrr/saving-plots-to-a-file-with-pdf-jpeg-and-png.html

png(
  file = here("doc/images/plot_ex_3_bonus.png"), # directory + file name
  width = 9, # width (in inches)
  height = 7 # height (in inches)
)

pirateplot(
  formula = ratings ~ imagery + drink + gender, # dependent ~ independent variables
  data = data_attitude_wine, # data frame
  main = "Bonus Plot Exercise 3", # plot title
  ylim = c(-25, 25), # y-axis: axis limits
  ylab = "", # y-axis: no label
  inf.method = "ci", # type of inference: 95% confidence interval
  inf.within = participant, # ID variable
  # theme settings
  pal = "espresso", # color palette: see piratepal("all")
  point.o = .5, # data points: opacity (0-1)
  point.cex = 1.3, # data points: size
  inf.f.o = .3, # 95% CI box filling: opacity (0-1)
  bean.b.o = .6, # bean border: opacity (0-1)
  bean.f.o = .6, # bean filling: opacity (0-1)
  cap.beans = TRUE, # bean densities are capped at the data limits
  bty = "n", # no box around the plot
  gl.col = "gray", # background line color (major and minor lines)
  gl.lwd = 1, # background line width
  gl.lty = 2 # background line type (dashed)
)

dev.off()

```

















